visitation_bias

### 2021-04-06 code/ms1_beta_diversity.Rmd
added management covariate to models

  envdat <- net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    #dplyr::select(bb.sp, plant.sp.abb, site) %>% #
    left_join(site_data) %>% # add site data
    dplyr::select(site, elev.mean, x, y, samples) %>%
    distinct() # return unique site-dates

@@

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits)
```

@@

# Join into single data frame for visualization
species_beta <- full_join(bb_beta, fl_beta) %>%
  left_join(site_data, by = c("site1" = "site")) %>% # add site data
  dplyr::select(site1, site2, FL, BB, # first for site 1
                elev1 = elev.mean, x1 = x, y1 = y) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, FL, BB, # then for site 2
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
  mutate(site1 = factor(site1), 
         site2 = factor(site2),
         elev.diff = elev1 - elev2,
         mean.elev = (elev1 + elev2)/2) %>%
  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
  filter(elev.diff >= 0) # For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes. Since no two sites have exactly the same elevation, this deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time. In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences. Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.

@@

# Add site data
beta_mit_data <- beta_site %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1 = elev.mean, x1 = x, y1 = y) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, 
                           elev.diff, mean.elev, 
                           x1, y1, x2, y2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
    metric = factor(metric, levels = c("WN", "ST", "OS", "S", "ST.h", "ST.l", 
                                       "ST.lh", "WN.repl", "WN.rich", 
                                       "OS.repl", "OS.rich")))
```

### Plot interaction beta-diversity against elevation difference and time difference
#### For elevation difference, we will "control" for time by only considering pairs separated by <= 5 days
#### For time difference, we will control for elevation by only comparing within site.
```{r}
# Set palette
# p <- c("OS" = "darkorange1", "ST" = "green3", 
#        "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
#        "ST.lh" = "deepskyblue", "WN" = "black")
# 
# 
# p <- c("OS" = "#66c2a5", "ST" = "#fc8d62", 
#        "ST.h" = "#8da0cb", "ST.l" = "#e78ac3", 
#        "ST.lh" = "#a6d854", "WN" = "black")

p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")

@@

survey_sample_diff <- survey %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

net_sample_diff <- net %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

mrm_survey_data <- species_beta %>%
  filter(group == "FL" & site1 != site2) %>%
  left_join(survey_sample_diff) %>%
  dplyr::select(site1, site2, beta, elev.diff, sample.diff) %>%
  mutate(elev.diff = scale(elev.diff), # gotta scale variables for MRM
         sample.diff = scale(sample.diff))

mrm_net_data <- species_beta %>%
  filter(group == "BB" & site1 != site2) %>%
  left_join(net_sample_diff)  %>%
  dplyr::select(site1, site2, beta, elev.diff, sample.diff) %>%
  mutate(elev.diff = scale(elev.diff),
         sample.diff = scale(sample.diff))

mrm_beta_data <- beta_mit_data %>%
  filter(site1 != site2) %>%
  left_join(net_sample_diff)  %>%
  dplyr::select(site1, site2, value, metric, elev.diff, sample.diff) %>%
  mutate(elev.diff = scale(elev.diff),
         sample.diff = scale(sample.diff))

mrm_fl <- MRM(beta ~ elev.diff + sample.diff, 
              data = mrm_survey_data,
              nperm = 1000,
              method = "logistic")

mrm_bb <- MRM(beta ~ elev.diff + sample.diff, 
              data = mrm_net_data,
              nperm = 1000,
              method = "logistic")

mrm_WN <- MRM(value ~ elev.diff + sample.diff, 
              data = filter(mrm_beta_data, metric == "WN"),
              nperm = 1000,
              method = "logistic")

mrm_ST <- MRM(value ~ elev.diff + sample.diff, 
              data = filter(mrm_beta_data, metric == "ST"),
              nperm = 1000,
              method = "logistic")

mrm_ST.l <- MRM(value ~ elev.diff + sample.diff, 
              data = filter(mrm_beta_data, metric == "ST.l"),
              nperm = 1000,
              method = "logistic")

mrm_ST.h <- MRM(value ~ elev.diff + sample.diff, 
              data = filter(mrm_beta_data, metric == "ST.h"),
              nperm = 1000,
              method = "logistic")

mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff, 
              data = filter(mrm_beta_data, metric == "ST.lh"),
              nperm = 1000,
              method = "logistic")

mrm_OS <- MRM(value ~ elev.diff + sample.diff, 
              data = filter(mrm_beta_data, metric == "OS"),
              nperm = 1000,
              method = "logistic")

@@

# Run GDM analysis
gdm_WN <- gdm(sitepair.WN, geo = TRUE)
summary(gdm_WN)

>>>>

  envdat <- net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    left_join(site_data) %>% # add site data
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() # return unique site-dates

@@

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) 
```

@@

# Join into single data frame for visualization
species_beta <- full_join(bb_beta, fl_beta) %>%
  left_join(site_data, by = c("site1" = "site")) %>% # add site data
  dplyr::select(site1, site2, FL, BB, # first for site 1
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, FL, BB, # then for site 2
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(site1 = factor(site1), 
         site2 = factor(site2),
         elev.diff = elev1 - elev2,
         mean.elev = (elev1 + elev2)/2, mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
  filter(elev.diff >= 0) # For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes. Since no two sites have exactly the same elevation, this deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time. In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences. Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.

@@

# Add site data
beta_mit_data <- beta_site %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, 
                           elev.diff, mean.elev, 
                           x1, y1, x2, y2, mgmt1, mgmt2, mgmt.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
    metric = factor(metric, levels = c("WN", "ST", "OS", "S", "ST.h", "ST.l", 
                                       "ST.lh", "WN.repl", "WN.rich", 
                                       "OS.repl", "OS.rich")))
```

### Plot interaction beta-diversity against elevation difference and time difference
#### For elevation difference, we will "control" for time by only considering pairs separated by <= 5 days
#### For time difference, we will control for elevation by only comparing within site.
```{r}
# Set palette
p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")

@@

survey_sample_diff <- survey %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

net_sample_diff <- net %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

mrm_survey_data <- species_beta %>%
  filter(group == "FL" & site1 != site2) %>%
  left_join(survey_sample_diff) %>%
  dplyr::select(site1, site2, beta, elev.diff, sample.diff, mgmt.diff) %>%
  mutate(elev.diff = scale(elev.diff), # gotta scale variables for MRM
         sample.diff = scale(sample.diff))

mrm_net_data <- species_beta %>%
  filter(group == "BB" & site1 != site2) %>%
  left_join(net_sample_diff)  %>%
  dplyr::select(site1, site2, beta, elev.diff, sample.diff, mgmt.diff) %>%
  mutate(elev.diff = scale(elev.diff),
         sample.diff = scale(sample.diff))

mrm_beta_data <- beta_mit_data %>%
  filter(site1 != site2) %>%
  left_join(net_sample_diff)  %>%
  dplyr::select(site1, site2, value, metric, elev.diff, sample.diff, mgmt.diff) %>%
  mutate(elev.diff = scale(elev.diff),
         sample.diff = scale(sample.diff))

mrm_fl <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_survey_data,
              nperm = 1000,
              method = "logistic")

mrm_bb <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_net_data,
              nperm = 1000,
              method = "logistic")

mrm_WN <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "WN"),
              nperm = 1000,
              method = "logistic")

mrm_ST <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST"),
              nperm = 1000,
              method = "logistic")

mrm_ST.l <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.l"),
              nperm = 1000,
              method = "logistic")

mrm_ST.h <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.h"),
              nperm = 1000,
              method = "logistic")

mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.lh"),
              nperm = 1000,
              method = "logistic")

mrm_OS <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "OS"),
              nperm = 1000,
              method = "logistic")

@@

# Run GDM analysis
gdm_WN <- gdm(sitepair.WN, geo = FALSE)
summary(gdm_WN)
gdm.varImp_WN <- gdm.varImp(sitepair.WN, geo = FALSE)

### 2021-04-06 code/ms2_bias_niche.Rmd
added management covariate to models

@@

# Plot intercept effect of tongue.diff
plot(pterm(niche_gam_21, 1))
```

>>>>

ggplot(niche_gam_21_pred, aes(elev.mean, mean.niche.overlap_trans)) +
  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
                  ymax = CI_upper_trans, ymin = CI_lower_trans),
              alpha = 0.3, inherit.aes = FALSE) +
  geom_line() +
  ylab("Niche overlap") +
  xlab("Elevation (m)") +
  guides(fill = FALSE) +
  facet_wrap(~tongue.pair) +
  theme_light()

@@

# Plot intercept effect of tongue.diff
pdf("../output/niche_gam_21_plot2.pdf", width = 5, height = 4)
plot(pterm(niche_gam_21, 1)) + labs(x = "Tongue pair", y = "Niche overlap") + theme_light()
dev.off()
```
